---
title: "Reprodicible Research: Assignment"
author: "Richa Agrawal"
date: "January 14, 2018"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## 1. Code for reading in the dataset and/or processing the data
```{r}
out.df<-read.csv("activity.csv",sep=",")
str(out.df)
out.df$date<-as.Date(out.df$date,format='%Y-%m-%d')
```

## 2. Histogram of the total number of steps taken each day
```{r}
stepsPerDay<-tapply(out.df$steps,out.df$date, sum, na.rm=TRUE)
hist(stepsPerDay, col="grey", xlab="Interval",main="Total Steps per day")
abline(v = mean(stepsPerDay), col="red",lwd=2)
abline(v = median(stepsPerDay), col="green",lwd=2)
```
## 3. Mean and median number of steps taken each day
```{r}
meanStepsByDate<-tapply(out.df$steps,out.df$date, mean, na.rm=TRUE)
medianStepsByDate<-tapply(out.df$steps,out.df$date, median,na.rm=TRUE)
cbind(meanStepsByDate,medianStepsByDate)
```
## 4. Time series plot of the average number of steps taken
```{r}
meanStepsByInterval<-tapply(out.df$steps,out.df$interval,mean,na.rm=TRUE)
df<-data.frame(meanStepsByInterval)

#plot(unique(out.df$interval),meanStepsByInterval,type="l",main = "Steps per Interval", xlab= "Interval", ylab= "Mean Steps", col="red" ,lwd=1)
plot(rownames(df),df$meanStepsByInterval,type="l",main = "Steps per Interval", xlab= "Interval", ylab= "Mean Steps", col="red" ,lwd=1)
```
## 5. The 5-minute interval that, on average, contains the maximum number of steps
```{r}
df[df$meanStepsByInterval==max(df$meanStepsByInterval),]
```
## 6. Strategy for imputing missing data
```{r}
new.df<-out.df
new.df$steps[which(is.na(new.df$steps))]<-mean(out.df$steps,na.rm=TRUE)

meanStepsByDateImp<-tapply(new.df$steps,new.df$date, mean)
medianStepsByDateImp<-tapply(new.df$steps,new.df$date, median,na.rm=TRUE)
cbind(meanStepsByDateImp,medianStepsByDateImp)
```
## 7. Histogram of the total number of steps taken each day after missing values are imputed
```{r}
stepsPerDayImp<-tapply(new.df$steps,new.df$date, sum, na.rm=TRUE)
hist(stepsPerDayImp, col="grey", xlab="Interval",main="Total Steps per day")
abline(v = mean(stepsPerDayImp), col="red",lwd=2)
abline(v = median(stepsPerDayImp), col="green",lwd=2, lty=2)
```
## 8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and 
weekends
```{r}
dframe<-out.df
dframe$day<-weekdays(dframe$date)
dframe$day<-ifelse(dframe$day == "Saturday" | dframe$day == "Sunday", "Weekend" , "Weekday")
dframe$day<-as.factor(dframe$day)
output<-tapply(dframe$steps, list(dframe$interval,dframe$day), sum, na.rm=TRUE)
output<-as.data.frame(output)

par(mfrow=c(2,1))
plot(rownames(output),output$Weekday, type='l', col="red", main="Interval vs SumStep on Weekdays", xlab = "Interval", ylab = "SumStep")
plot(rownames(output),output$Weekend, type='l', col="blue", main="Interval vs SumStep on Weekends", xlab = "Interval", ylab = "SumStep")
```
## 9. All of the R code needed to reproduce the results (numbers, plots, etc.) in the report
See all the figures and code above from 1-8.
